---
title: "Pratical Machine Learning"
author: "Weihua Shen"
date: "2019/11/3"
output: html_document
---

## Background 

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.

In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Library and loading data
```{r}
library(caret)
library(rattle)
TrainData <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"),header=TRUE)
dim(TrainData)
TestData <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"),header=TRUE)
dim(TestData)
str(TrainData)
# Here we get the indexes of the columns having at least 90% of NA or blank values on the training dataset
indColToRemove <- which(colSums(is.na(TrainData) |TrainData=="")>0.9*dim(TrainData)[1]) 
TrainDataClean <- TrainData[,-indColToRemove]
TrainDataClean <- TrainDataClean[,-c(1:7)]
dim(TrainDataClean)
# We do the same for the test set
indColToRemove <- which(colSums(is.na(TestData) |TestData=="")>0.9*dim(TestData)[1]) 
TestDataClean <- TestData[,-indColToRemove]
TestDataClean <- TestDataClean[,-1]
dim(TestDataClean)
str(TestDataClean)
# Here we create a partition of the traning data set 
set.seed(12345)
inTrain1 <- createDataPartition(TrainDataClean$classe, p=0.75, list=FALSE)
Train1 <- TrainDataClean[inTrain1,]
Test1 <- TrainDataClean[-inTrain1,]
dim(Train1)
dim(Test1)
```

## Training with decision tree
```{r}
trControl <- trainControl(method="cv", number=5)
model_CT <- train(classe~., data=Train1, method="rpart", trControl=trControl)
#print(model_CT)
fancyRpartPlot(model_CT$finalModel)
trainpred <- predict(model_CT,newdata=Test1)
confMatCT <- confusionMatrix(Test1$classe,trainpred)
# display confusion matrix and model accuracy
confMatCT$table
confMatCT$overall[1]
```

## Random Forest
```{r}
model_RF <- train(classe~., data=Train1, method="rf", trControl=trControl, verbose=FALSE)
print(model_RF)
plot(model_RF,main="Accuracy of Random forest model by number of predictors")
trainpred <- predict(model_RF,newdata=Test1)
confMatRF <- confusionMatrix(Test1$classe,trainpred)
# display confusion matrix and model accuracy
confMatRF$table
confMatRF$overall[1]
names(model_RF$finalModel)
model_RF$finalModel$classes
plot(model_RF$finalModel,main="Model error of Random forest model by number of trees")
# Compute the variable importance 
MostImpVars <- varImp(model_RF)
MostImpVars
```

## Train with gradient boosting method
```{r}
model_GBM <- train(classe~., data=Train1, method="gbm", trControl=trControl, verbose=FALSE)
print(model_GBM)
plot(model_GBM)
trainpred <- predict(model_GBM,newdata=Test1)
confMatGBM <- confusionMatrix(Test1$classe,trainpred)
confMatGBM$table
confMatGBM$overall[1]
```

## Conclusion
The random forest seems the best one. And here we use random forest trained model to predict
```{r}
FinalTestPred <- predict(model_RF,newdata=TestDataClean)
FinalTestPred
```

